\bf Classification and Regression on Random Dot Product Graphs

### 2025-10-17 draft.Rmd
finished cleaning up proof of theorem 3, still working on proof of theorem 4

doMC::registerDoMC(parallel::detectCores())
```

@@

```{r angle-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
n.sigma <- 1

@@

angle.reg.df %>% 
  dplyr::filter(n >= 128) %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(mse.theta = mean(mse),
                   se.theta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
                   se.m = sd(mse.m) / sqrt(dplyr::n()) * n.sigma,
                   mse.m = mean(mse.m),
                   se.a = sd(mse.a) / sqrt(dplyr::n()) * n.sigma,
                   mse.a = mean(mse.a)) %>% 
  ggplot() + 
  theme_bw() + 
  scale_colour_brewer(palette = 'Set1') +
  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor') +
  geom_line(aes(x = n, y = mse.theta, colour = 'estimated \nangle \n')) + 
  geom_errorbar(aes(x = n, 
                    ymin = mse.theta - se.theta, 
                    ymax = mse.theta + se.theta,
                    colour = 'estimated \nangle \n'), 
                width = .1) + 
  # geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
  # geom_errorbar(aes(x = n,
  #                   ymin = mse.m - se.m,
  #                   ymax = mse.m + se.m,
  #                   colour = 'modularity'), 
  #               width = .1) + 
  geom_line(aes(x = n, y = mse.a, colour = 'graph \nstatistics')) + 
  geom_errorbar(aes(x = n,
                    ymin = mse.a - se.a,
                    ymax = mse.a + se.a,
                    colour = 'graph \nstatistics'), 
                width = .1) + 
  scale_x_log10(breaks = sort(n.vec))
```

```{r theta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
angle.reg.df %>% 
  dplyr::filter(n >= 128) %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(theta.se = sd(theta.mse) / sqrt(dplyr::n()) * n.sigma,
                   theta.mse = mean(theta.mse)) %>% 
  ggplot() + 
  geom_line(aes(x = n, y = theta.mse)) + 
  geom_errorbar(aes(x = n, ymin = theta.mse - theta.se, ymax = theta.mse + theta.se),
                width = .1) + 
  labs(x = 'Expected vertices per graph', y = expression('MSE('~hat(theta)~')')) + 
  theme_bw() + 
  scale_y_log10() + 
  scale_x_log10(breaks = sort(n.vec))
```

```{r beta-regression-sim, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}

@@

beta.reg.df %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(mse.beta = mean(mse),
                   se.beta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
                   se.d = sd(mse.d) / sqrt(dplyr::n()) * n.sigma,
                   mse.d = mean(mse.d),
                   se.t = sd(mse.t) / sqrt(dplyr::n()) * n.sigma,
                   mse.t = mean(mse.t)) %>% 
  ggplot() + 
  geom_line(aes(x = n, y = mse.beta, colour = 'estimated \nbeta \nparameters\n')) + 
  geom_errorbar(aes(x = n, 
                    ymin = mse.beta - se.beta, 
                    ymax = mse.beta + se.beta,
                    colour = 'estimated \nbeta \nparameters\n'),
                width = .1) + 
  geom_line(aes(x = n, y = mse.d, colour = 'graph \nstatistics')) + 
  geom_errorbar(aes(x = n,
                    ymin = mse.d - se.d,
                    ymax = mse.d + se.d,
                    colour = 'graph \nstatistics'),
                width = .1) + 
  # geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
  # geom_errorbar(aes(x = n,
  #                   ymin = mse.t - se.t,
  #                   ymax = mse.t + se.t,
  #                   colour = 'transitivity'),
  #               width = .1) +
  theme_bw() + 
  scale_colour_brewer(palette = 'Set1') +
  scale_x_log10(breaks = sort(n.vec)) + 
  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor')
```

```{r beta-mse, fig.width = 6, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.'}
beta.reg.df %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(a.se = sd(a.mse) / sqrt(dplyr::n()) * n.sigma,
                   b.se = sd(b.mse) / sqrt(dplyr::n()) * n.sigma,
                   a.mse = mean(a.mse),
                   b.mse = mean(b.mse)) %>% 
  ggplot() + 
  # scale_y_log10() + 
  scale_x_log10() + 
  geom_line(aes(x = n, y = a.mse, colour = 'a')) + 
  geom_errorbar(aes(x = n, ymin = a.mse - a.se, ymax = a.mse + a.se, colour = 'a'),
                width = .1) + 
  geom_line(aes(x = n, y = b.mse, colour = 'b')) + 
  geom_errorbar(aes(x = n, ymin = b.mse - b.se, ymax = b.mse + b.se, colour = 'b'),
                width = .1) + 
  theme_bw() + 
  labs(x = 'number of vertices', y = 'MSE of Beta \nparameter estimates', 
       colour = 'Beta \ndistribution \nparameter') + 
  scale_colour_brewer(palette = 'Set1') +
  scale_y_log10() + 
  scale_x_log10(breaks = sort(n.vec))
```

@@

In this section, we apply \ref{alg:mslm-fit} to real data to see if we can extract estimates from the latent space that correspond to response variables. 
In the first example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
This dataset consists of two graphs representing two networks of neurons, one for each hemisphere of the *Drosophila* brain. 
In these graphs, each vertex is a neuron, and the labels correspond to one of four neuron types (Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons). 
The number of neurons in each hemisphere is not equal (209 in the left hemisphere and 213 in the right hemisphere).
The resulting graphs are illustrated in figure \ref{fig:mbconnectome-graph}.

@@

```{r load-hcp}
load('~/Downloads/Data/HCP/HCP-A/DTI/Data.RData')
z <- dti.info$hemisphere %>% 
  as.factor()
rois <- dti.info$aparc_aseg_region.names.short

@@

```{r}
pairwise.angles.df <- readr::read_csv('~/dev/multilayer-rdpg/pairwise-angles.csv')

@@

\begin{proof}[Proof of Theorem \ref{thm:theta-consistency}]
By theorem \ref{thm:bezier-consistency}, $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$. 
Each element of $T$ is $T_{ir} = \binom{R}{r} t_i^r (1 - t_i)^{R-r}$, and similarly $\hat{T}_{ir} = \binom{R}{r} \hat{t}_i^r (1 - \hat{t}_i)^{R-r}$. 
This is unique for the class of Bezier polynomials restricted by algorithm 3. 
Then $\sum_i^n \|t_i - \hat{t}_i\| \stackrel{p}{\to} 0$. 
\end{proof}

>>>>

doMC::registerDoMC(parallel::detectCores() - 2)
```

@@

```{r angle-regression-sim, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 1. The errorbars are the standard error.'}
n.sigma <- 1

@@

angle.reg.df %>% 
  dplyr::filter(n >= 128) %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(mse.theta = mean(mse),
                   se.theta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
                   se.m = sd(mse.m) / sqrt(dplyr::n()) * n.sigma,
                   mse.m = mean(mse.m),
                   se.a = sd(mse.a) / sqrt(dplyr::n()) * n.sigma,
                   mse.a = mean(mse.a)) %>% 
  ggplot() + 
  theme_bw() + 
  scale_colour_brewer(palette = 'Set1') +
  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor') +
  geom_line(aes(x = n, y = mse.theta, colour = 'Estimated \nangle \n')) + 
  geom_errorbar(aes(x = n, 
                    ymin = mse.theta - se.theta, 
                    ymax = mse.theta + se.theta,
                    colour = 'Estimated \nangle \n'), 
                width = .1) + 
  # geom_line(aes(x = n, y = mse.m, colour = 'modularity')) + 
  # geom_errorbar(aes(x = n,
  #                   ymin = mse.m - se.m,
  #                   ymax = mse.m + se.m,
  #                   colour = 'modularity'), 
  #               width = .1) + 
  geom_line(aes(x = n, y = mse.a, colour = 'Graph \nstatistics')) + 
  geom_errorbar(aes(x = n,
                    ymin = mse.a - se.a,
                    ymax = mse.a + se.a,
                    colour = 'Graph \nstatistics'), 
                width = .1) + 
  scale_x_log10(breaks = sort(n.vec))
```

```{r theta-mse, fig.width = 4, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated angles in scenario 1 over 50 simulations. The errorbars are the standard error.'}
angle.reg.df %>% 
  dplyr::filter(n >= 128) %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(theta.se = sd(theta.mse) / sqrt(dplyr::n()) * n.sigma,
                   theta.mse = mean(theta.mse)) %>% 
  ggplot() + 
  geom_line(aes(x = n, y = theta.mse)) + 
  geom_errorbar(aes(x = n, ymin = theta.mse - theta.se, ymax = theta.mse + theta.se),
                width = .1) + 
  labs(x = 'Expected vertices per graph', y = expression('MSE('~hat(theta)~')')) + 
  theme_bw() + 
  scale_y_log10() + 
  scale_x_log10(breaks = sort(n.vec))
```

```{r beta-regression-sim, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE over 50 simulations in scenario 2. The errorbars are the standard error.'}

@@

beta.reg.df %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(mse.beta = mean(mse),
                   se.beta = sd(mse) / sqrt(dplyr::n()) * n.sigma,
                   se.d = sd(mse.d) / sqrt(dplyr::n()) * n.sigma,
                   mse.d = mean(mse.d),
                   se.t = sd(mse.t) / sqrt(dplyr::n()) * n.sigma,
                   mse.t = mean(mse.t)) %>% 
  ggplot() + 
  geom_line(aes(x = n, y = mse.beta, colour = 'Estimated \nbeta \nparameters\n')) + 
  geom_errorbar(aes(x = n, 
                    ymin = mse.beta - se.beta, 
                    ymax = mse.beta + se.beta,
                    colour = 'Estimated \nbeta \nparameters\n'),
                width = .1) + 
  geom_line(aes(x = n, y = mse.d, colour = 'Graph \nstatistics')) + 
  geom_errorbar(aes(x = n,
                    ymin = mse.d - se.d,
                    ymax = mse.d + se.d,
                    colour = 'Graph \nstatistics'),
                width = .1) + 
  # geom_line(aes(x = n, y = mse.t, colour = 'transitivity')) + 
  # geom_errorbar(aes(x = n,
  #                   ymin = mse.t - se.t,
  #                   ymax = mse.t + se.t,
  #                   colour = 'transitivity'),
  #               width = .1) +
  theme_bw() + 
  scale_colour_brewer(palette = 'Set1') +
  scale_x_log10(breaks = sort(n.vec)) + 
  labs(x = 'Expected vertices per graph', y = 'MSE', colour = 'Predictor')
```

```{r beta-mse, fig.width = 5, fig.height = 2, cache = FALSE, fig.cap = 'The average MSE of the estimated beta parameters, $(a, b)$, in scenario 2 over 50 simulations. The errorbars are the standard error.'}
beta.reg.df %>% 
  dplyr::group_by(n) %>% 
  dplyr::summarise(a.se = sd(a.mse) / sqrt(dplyr::n()) * n.sigma,
                   b.se = sd(b.mse) / sqrt(dplyr::n()) * n.sigma,
                   a.mse = mean(a.mse),
                   b.mse = mean(b.mse)) %>% 
  ggplot() + 
  # scale_y_log10() + 
  scale_x_log10() + 
  geom_line(aes(x = n, y = a.mse, colour = 'a')) + 
  geom_errorbar(aes(x = n, ymin = a.mse - a.se, ymax = a.mse + a.se, colour = 'a'),
                width = .1) + 
  geom_line(aes(x = n, y = b.mse, colour = 'b')) + 
  geom_errorbar(aes(x = n, ymin = b.mse - b.se, ymax = b.mse + b.se, colour = 'b'),
                width = .1) + 
  theme_bw() + 
  labs(x = 'number of vertices', y = 'MSE of Beta \nparameter estimates', 
       colour = 'Beta \ndistribution \nparameter') + 
  scale_colour_brewer(palette = 'Set1') +
  scale_y_log10() + 
  scale_x_log10(breaks = sort(n.vec))
```

@@

In this section, we apply algorithm \ref{alg:mslm-fit} to real data to see if we can extract estimates from the latent space that correspond to response variables. 
In the first example, we analyzed the larval *Drosophila* mushroom body connectome \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2020estimation}. 
This dataset consists of two graphs representing two networks of neurons, one for each hemisphere of the *Drosophila* brain. 
In these graphs, each vertex is a neuron, and the labels correspond to one of four neuron types (Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons). 
The number of neurons in each hemisphere is not equal (209 in the left hemisphere and 213 in the right hemisphere).
The resulting graphs are illustrated in figure \ref{fig:mbconnectome-graph}.

@@

```{r load-hcp}
load('data/Data.RData')
z <- dti.info$hemisphere %>% 
  as.factor()
rois <- dti.info$aparc_aseg_region.names.short

@@

```{r}
pairwise.angles.df <- readr::read_csv('data/pairwise-angles.csv')

@@

\begin{proof}[Proof of Theorem \ref{thm:theta-consistency}]
By theorem \ref{thm:bezier-consistency}, $\|T - \hat{T}\|_{2, \infty} \stackrel{p}{\to} 0$. 
Each element of $T$ is $T_{ir} = \binom{R}{r} t_i^r (1 - t_i)^{R-r}$, and similarly $\hat{T}_{ir} = \binom{R}{r} \hat{t}_i^r (1 - \hat{t}_i)^{R-r}$. 
This is unique for the class of Bezier polynomials restricted by algorithm 3. 
Then $\sum_i^n |t_i - \hat{t}_i| \stackrel{p}{\to} 0$, so each $|t_i - \hat{t}_i| \stackrel{p}{\to} 0$. 
By construction, the MLE $\hat{\theta}$ and the quasi-MLE $\tilde{\theta}$ are computed via the same function, just on different samples, i.e., $\hat{\theta} = h(t_1, ..., t_n)$ and $\tilde{\theta} = h(\hat{t}_1, ..., \hat{t}_n)$. 
Then since the probability distribution function of each $t_i$ is smooth, $h$ is continuous on the support. 
Therefore, if each $\hat{t}_i \stackrel{p}{\to} t$, then $h(\hat{t}_1, ..., \hat{t}_n) \stackrel{p}{\to} h(t_1, ..., t_n)$, resulting in $\tilde{\theta} \stackrel{p}{\to} \hat{\theta} \stackrel{p}{\to} \theta$.

\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:supervised}]
\end{proof}

### 2025-10-17 draft.tex
finished cleaning up proof of theorem 3, still working on proof of theorem 4

In this section, we apply \ref{alg:mslm-fit} to real data to see if we
can extract estimates from the latent space that correspond to response
variables. In the first example, we analyzed the larval
\emph{Drosophila} mushroom body connectome \citep{Eichler141762}, which
has been studied as a GRDPG by \citet{athreya2020estimation}. This
dataset consists of two graphs representing two networks of neurons, one
for each hemisphere of the \emph{Drosophila} brain. In these graphs,
each vertex is a neuron, and the labels correspond to one of four neuron
types (Kenyon Cells, Input Neurons, Output Neurons, and Projection
Neurons). The number of neurons in each hemisphere is not equal (209 in
the left hemisphere and 213 in the right hemisphere). The resulting
graphs are illustrated in figure \ref{fig:mbconnectome-graph}.

>>>>

In this section, we apply algorithm \ref{alg:mslm-fit} to real data to
see if we can extract estimates from the latent space that correspond to
response variables. In the first example, we analyzed the larval
\emph{Drosophila} mushroom body connectome \citep{Eichler141762}, which
has been studied as a GRDPG by \citet{athreya2020estimation}. This
dataset consists of two graphs representing two networks of neurons, one
for each hemisphere of the \emph{Drosophila} brain. In these graphs,
each vertex is a neuron, and the labels correspond to one of four neuron
types (Kenyon Cells, Input Neurons, Output Neurons, and Projection
Neurons). The number of neurons in each hemisphere is not equal (209 in
the left hemisphere and 213 in the right hemisphere). The resulting
graphs are illustrated in figure \ref{fig:mbconnectome-graph}.
